Attract and kill: Spinosad containing spheres to control onion maggot (\textit{Delia antiqua})

### 2019-11-11 manuscript/pest-management-science/main.tex

Onion maggot (\textit{Delia antiqua}) is a pest of onions worldwide. Current means of managing this pest rely heavily on prophylatic insecticide treatments at planting. These options may not be viable in organic production systems or situations where insecticide-resistant populations occur. Here we explore the efficacy of an attract and kill strategy for control of \textit{D. antiqua} evaluating the ability of attractive, spinosad containing spheres to kill adult \textit{D. antiqua} and reduce crop losses. Spinosad containing spheres were able to consistently kill \textit{D. antiqua} adults over the course of the field season (mortality range: between 49\% and 59\% on average). Pairing spinosad spheres with Delia Lure increased efficacy by 72\% compared with the attract and kill strategy alone. Performance of this attract and kill strategy also can reduce damage by \textit{D. antiqua}  larvae in the field, but it did not achieve a level of control comparable to the level provided by a conventional insecticide treatment. Implementation of this attract and kill strategy could be a valuable tool in situations where conventional pesticides are either not available or desired, where additional control techniques are needed, or to provide a season-long option for control of \textit{D. antiqua} populations.

>>>>

BACKGROUND: Onion maggot (\textit{Delia antiqua}) is a pest of onions worldwide. Current means of managing this pest rely heavily on prophylatic insecticide treatments at planting. These options may not be viable in organic production systems or situations where insecticide-resistant populations occur. Here we explore the efficacy of an attract and kill strategy for control of \textit{D. antiqua} evaluating the ability of attractive, spinosad containing spheres to kill adult \textit{D. antiqua} and reduce crop losses.

RESULTS: Spinosad containing spheres were able to consistently kill \textit{D. antiqua} adults over the course of the field season (mortality range: between 49\% and 59\% on average). Pairing spinosad spheres with Delia Lure increased efficacy by 72\% compared with the spheres alone. Performance of this attract and kill strategy also can reduce damage by \textit{D. antiqua}  larvae in the field, but it did not achieve a level of control comparable to the level provided by a conventional insecticide treatment.

CONCLUSION: Implementation of this attract and kill strategy could be a valuable tool in situations where conventional pesticides are either not available or desired, where additional control techniques are needed, or to provide a season-long option for control of \textit{D. antiqua} populations.
